Uneven Investments in Toronto's Wards

### 2024-09-27 paper/paper.qmd
updated paper.pdf with final changes
- updated position on figures
- small grammatical changes

-   Per Capita Funding: Total capital funding allocated to a ward divided by its population.

-   Median Household Income: The median income of households within a ward.

-   Low-Income Prevalence: The percentage of households classified as low-income in a ward.

-   Educational Attainment: The percentage of residents with a bachelor's degree or higher.

-   Visible Minority Population: The percentage of residents who identify as visible minorities.

@@

```{r Figure 1, echo=FALSE, message=FALSE, warning=FALSE}
#| label: fig-percapitafunding
#| fig-cap: Bar graph showing Per Capita Capital Funding by Ward

# Wards by Per Capita Funding
ggplot(merged_data, aes(x = reorder(`Ward`, per_capita_funding), y = per_capita_funding)) +
  geom_bar(stat = "identity", fill = "#0072B2") +
  labs(
    title = "Wards by Per Capita Funding",
    x = "Ward Name - Number",
    y = "Per Capita Funding ($CAD)"
  ) +
  scale_y_continuous(labels = dollar_format()) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 14, hjust = 0.5),
    axis.text.x = element_text(angle = 45, hjust = 1, size = 10),
    axis.title = element_text(size = 14),
    axis.text.y = element_text(size = 10)
  ) +
  coord_flip()
```

@@

The average income across wards is \$84,000, meaning half of the wards have median incomes at or below this amount. Higher income wards will be categorized as one with a median household income above \$84,000. A positive relationship was observed between per capita funding and median household income as wards above the citywide average \$84,000 generally received more funding per capita. Specifically, among the top five wards in per capita funding, four have median incomes above \$84,000. However, among the bottom five wards, four have median incomes below \$84,000.

More specifically, Ward 2, with a high median income of \$100,000, received the second-highest per capita funding at \$4,166.43. In contrast, Ward 22, which has a lower median income of \$77,000, received the second-lowest per capita funding at just \$192.44. This pattern suggests that wealthier wards benefit more from capital investments.

```{r Figure 2, echo=FALSE, message=FALSE, warning=FALSE}
#| label: fig-householdincome
#| fig-cap: Scatter plot illustrating the relationship between Per Capita Funding and Median Household Income

@@

```{r Figure 3, echo=FALSE, message=FALSE, warning=FALSE}
#| label: fig-lowincome
#| fig-cap: Per capita funding and low income prevalence for top and bottom 5 wards excluding ward 13

@@

We will also classify wards where at least 33% of residents hold a bachelor's degree or higher as having high educational attainment. There is a clear positive relationship between educational attainment and per capita funding as wards with higher education levels tend to receive more funding. More specifically, Ward 10, with 60.97% of residents holding a bachelor’s degree, received \$4,127.10 per capita, while Ward 11, with 56.07%, received \$1,637.81 per capita. In contrast, Ward 7, where only 14.67% of residents hold a degree, received just \$852.05 per capita.

```{r Figure 4, echo=FALSE, message=FALSE, warning=FALSE}
#| label: fig-education
#| fig-cap: Scatter plot illustrating the relationship between Per Capita Funding and Education Levels

@@

```{r Figure 5, echo=FALSE, message=FALSE, warning=FALSE}
#| label: fig-minority
#| fig-cap: Scatter plot illustrating the relationship between Per Capita Funding and Minority Population

>>>>

-   `Per Capita Funding`: Total capital funding allocated to a ward divided by its population.

-   `Median Household Income`: The median income of households within a ward.

-   `Low-Income Prevalence`: The percentage of households classified as low-income in a ward.

-   `Educational Attainment`: The percentage of residents with a bachelor's degree or higher.

-   `Visible Minority Population`: The percentage of residents who identify as visible minorities.

@@

```{r Figure 1, echo=FALSE, message=FALSE, warning=FALSE}
#| label: fig-percapitafunding
#| fig-cap: Bar graph showing Per Capita Capital Funding by Ward
#| fig-pos: H

# Wards by Per Capita Funding
ggplot(merged_data, aes(x = reorder(`Ward`, per_capita_funding), y = per_capita_funding)) +
  geom_bar(stat = "identity", fill = "#0072B2") +
  labs(
    title = "Wards by Per Capita Funding",
    x = "Ward Name - Number",
    y = "Per Capita Funding ($CAD)"
  ) +
  scale_y_continuous(labels = dollar_format()) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 12, hjust = 0.5),
    axis.text.x = element_text(angle = 45, hjust = 1, size = 8),
    axis.title = element_text(size = 10),
    axis.text.y = element_text(size = 8)
  ) +
  coord_flip()
```

@@

The average income across wards is \$84,000, meaning half of the wards have median incomes at or below this amount. Higher income wards will be categorized as one with a median household income above \$84,000.

A positive relationship was observed between per capita funding and median household income as wards above the citywide average \$84,000 generally received more funding per capita. Specifically, among the top five wards in per capita funding, four have median incomes above \$84,000. However, among the bottom five wards, four have median incomes below \$84,000. Ward 2, with a high median income of \$100,000, received the second-highest per capita funding at \$4,166.43. In contrast, Ward 22, which has a lower median income of \$77,000, received the second-lowest per capita funding at just \$192.44. This pattern suggests that wealthier wards benefit more from capital investments.

```{r Figure 2, echo=FALSE, message=FALSE, warning=FALSE}
#| label: fig-householdincome
#| fig-cap: Scatter plot illustrating the relationship between Per Capita Funding and Median Household Income
#| fig-pos: H

@@

```{r Figure 3, echo=FALSE, message=FALSE, warning=FALSE}
#| label: fig-lowincome
#| fig-cap: Per capita funding and low income prevalence for top and bottom 5 wards excluding ward 13 
#| fig-pos: H

@@

We will classify wards where at least 33% of residents hold a bachelor's degree or higher, as having high educational attainment. There is a clear positive relationship between educational attainment and per capita funding as wards with higher education levels tend to receive more funding. More specifically, Ward 10, with 60.97% of residents holding a bachelor’s degree, received \$4,127.10 per capita, while Ward 11, with 56.07%, received \$1,637.81 per capita. In contrast, Ward 7, where only 14.67% of residents hold a degree, received just \$852.05 per capita.

```{r Figure 4, echo=FALSE, message=FALSE, warning=FALSE}
#| label: fig-education
#| fig-cap: Scatter plot illustrating the relationship between Per Capita Funding and Education Levels
#| fig-pos: H

@@

```{r Figure 5, echo=FALSE, message=FALSE, warning=FALSE}
#| label: fig-minority
#| fig-cap: Scatter plot illustrating the relationship between Per Capita Funding and Minority Population
#| fig-pos: H
